\documentclass[11pt, a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{parskip}

\title{Supplementary Results: Naive Baseline Comparison\\
\large Hypoglycaemia Event Detection -- T1DATA Cohort (36 Patients)}
\author{Robin van den Hoek}
\date{\today}

\begin{document}
\maketitle

\section*{Overview}

This document reports the performance of two naive rule-based detectors alongside the trained
event classifiers. The naive baselines require no learning and serve as a lower bound: a trained
model that does not outperform them has not extracted any useful information from the data.

All metrics are cohort means over 36 patients, evaluated on the held-out test split.

\section*{Naive Baseline Definitions}

\textbf{Naive Last Value.}
If the most recent glucose reading in the lookback window is below 70~mg/dL, an event is
predicted. This rule fires only when the patient is already in hypoglycaemia. It is highly
precise but misses the large majority of events because it cannot anticipate future drops.

\textbf{Naive Mean.}
If the average glucose over the full 2-hour lookback window is below 70~mg/dL, an event is
predicted. Because the average must fall below the threshold, this rule fires very rarely and
misses almost everything.

\section*{Task A: Will a hypoglycaemia event occur anywhere in the next 60 minutes?}

\begin{table}[h]
\centering
\caption{Task A results. F2 weights recall twice as heavily as precision.}
\begin{tabular}{lrrrr}
\toprule
Model & Precision & Recall & F1 & F2 \\
\midrule
Naive Last Value        & 0.857 & 0.326 & 0.473 & 0.373 \\
Naive Mean              & 0.396 & 0.041 & 0.074 & 0.050 \\
\midrule
LSTM Event Classifier   & 0.237 & 0.693 & 0.334 & 0.463 \\
PatchTST Event Classifier & 0.054 & 0.615 & 0.095 & 0.178 \\
\bottomrule
\end{tabular}
\end{table}

The Naive Last Value achieves a higher F1 (0.473) than the LSTM classifier (0.334) because it
operates at a very different precision--recall trade-off: it fires only when glucose is already
below 70~mg/dL, which makes it very precise but nearly useless for early warning. The LSTM
classifier has a substantially higher recall (0.693 vs.\ 0.326), meaning it catches more than
twice as many events at an earlier point in time. On F2, which weights recall more heavily,
the LSTM classifier (0.463) outperforms the naive rule (0.373). The PatchTST classifier does
not outperform the Naive Last Value on any metric.

\section*{Task B: Will a hypoglycaemia event occur at approximately 60 minutes (tolerance $\pm$15 min)?}

\begin{table}[h]
\centering
\caption{Task B results. The positive label is 1 if the minimum glucose in the window [45, 75]~min ahead is below 70~mg/dL.}
\begin{tabular}{lrrrr}
\toprule
Model & Precision & Recall & F1 & F2 \\
\midrule
Naive Last Value              & 0.216 & 0.113 & 0.149 & 0.125 \\
Naive Mean                    & 0.115 & 0.016 & 0.029 & 0.020 \\
\midrule
LSTM Point-Event Classifier   & 0.088 & 0.525 & 0.143 & 0.237 \\
PatchTST Point-Event Classifier & 0.036 & 0.537 & 0.064 & 0.124 \\
\bottomrule
\end{tabular}
\end{table}

Task B is a harder and more specific question. The LSTM point-event classifier barely exceeds
the Naive Last Value on F1 (0.143 vs.\ 0.149) and has substantially lower precision (0.088
vs.\ 0.216). On F2, the LSTM (0.237) outperforms the naive rule (0.125) due to its higher
recall. The PatchTST point-event classifier does not outperform the Naive Last Value on any
metric. These results indicate that predicting the exact 60-minute horizon is a considerably
harder problem than predicting whether any event will occur in the next hour.

\section*{Note on Inclusion in Thesis}

The naive baselines are reported here for completeness and supervisor reference. Following
discussion with the supervisor, they are not included in the main thesis, as the focus of the
ablation study is on the comparison between forecasting-derived detectors and trained binary
classifiers rather than on rule-based baselines.

\end{document}
